Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-20 drafts/hybrid_zone.tex
edits throughout

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{color}
\usepackage{amssymb}
\usepackage{setspace}

@@

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.  Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}. 
%The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

@@

Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. Our goal is to understand how selection against hybrids interacts with dispersal and recombination interact to produce observed patterns ancestry around selected sites in recently established hybrid zones. In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness. In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and  patterns of introgression can provide information about selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}. 
\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}

@@

We model selection as a simple case of under dominance at a single locus.  Specifically, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes. \yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those degenerated under complex epistatic systems}.

Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at time $\tau$ generations in the past (Figure~\ref{Fig:schematic}). 
The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus). 
For every other locus, we keep track of $a$, the identity at $L$, across time and space. If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise. Recombination between two loci is modeled as a poisson process with rate $r$. The identity $a$ at locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.

\begin{figure}
\includegraphics{figs/BM_schematic}
\caption{A cartoon representation of a haplotype under the model. The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day ($\tau=0$). The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A segregates at lower frequency for $x>0$. Here, because of negative selection, $L_B$ migrates rapidly back to $x>0$ from a position $x<0$. Red dots represent recombination events along the entire haplotype, and lineages that recombine off the same haplotype as $L_B$ are represented by colored lines. (here, the initial haplotype is broken up into 3 chunks over 100 generations). Subsequent recombination events can cause these chunks to change which allele they are linked to at the selected site, and here yellow lineages are linked to an allele of ancestry $A$ at the selected locus, and blue lineages are linked to a selected allele of ancestry $B$. From right to left, the chunks at $\tau=100$ are of ancestry $(A, B, B)$. }\label{Fig:schematic}
\end{figure}
%As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone. This means that recombination events that occur away from the center of the zone do not result in a change in $a$.

@@

Consider the collection of $A$ alleles found at location $x$.
The expected number of such alleles that had a parent at location $y$ in the previous generation is proportional to the number of  $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$. In other words, lineages move as a random walk determined by the dispersal kernel,
but biased towards locations where the selected allele they carry is at higher frequency.

\alisa{Maybe describe, in a few words, how you're constructing the model that is in the Appendix.}
Making the same assumptions underlying equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill \alisa{better word?} on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
Therefore, roughly, lineages on 
then roughly, lineages on ``their own'' side wander randomly,
while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
back towards the side where they are more common. \alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
Lineages of $B$ alleles do the same thing, of course, but in the opposite direction. \alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}

\alisa{Where did the line numbers go??}
\paragraph{Lineages at linked loci}
The behavior of a lineage at a linked locus is similar to a selected locus,
except that when in heterozygotes the lineage may move between backgrounds:
if we follow back through time the lineage of a locus we find linked to an $A$ allele, 
it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
but if it spends time in the clinal region where recombination may happen,
it may be found to have been inherited from a $B$-carrying individual,
whose ancestors will tend to be more from the right.

@@

\paragraph{PDE for linked clines}
Clines at linked sites are given by
the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
We denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation. \alisa{Is the above equation true? Shouldn't it depend on $r$?}
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            + r (1-p(x,t))(q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} \\
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$. \alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
We will have more use for the differential operators on the right-hand sides of these equations,
so write these also as 
$$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$$
and 
$$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$$.
\alisa{A few words about what $G$ is?}

@@

\paragraph{Haplotype lengths}
We can also find the probability of seeing an entire block of genome (haplotype)
of a single ancestry at a given location and time.
As in \citet{Sedghifar2015},
to do this we need only keep track of both lineages when a recombination occurs within a block (Figure~\ref{Fig:schematic}),
so that lineages will act like a particular class of labeled, branching diffusions,
where the total branching rate is conserved. We will not consider subsequent coalescence.
The general description of the process is as follows:
The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
with recombination distance $r$ equal to the rate at which the segment recombines away from the selected site.
(If the selected site lies inside the segment, $r=0$.)
Additionally, at rate equal to the genetic length of the segment,
a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
each of which proceeds as before, independently.

\alisa{indeed, regions surrounding the selected locus have a hard time introgressing (Figure~\ref{Fig:resistanceToIntrogression100g})}

\paragraph{General haplotype probabilities}
To quantify haplotype abundances,
we'd like to know the probability $g_A(x,t;a,b)$ that the entire segment of genome between positions $a$ and $b$, relative to the selected site,
in an individual sampled at spatial location $x$ at time $t$ after the cline has formed is of ancestry $A$.
This is similar to equation \eqref{eqn:q_pde} with a term added for branching, so that 
\plr{write this equation more compactly?}
\begin{align}
    \begin{aligned} \label{eqn:g_pde}
        \dot g_A(x,t;a,b) 
            &= G_{r(a,b)}^A g_A(x,t;a,b) 
            - (b-a) g_A(x,t;a,b) 
            \\ {} & \qquad 
            + p(x,t) \int_a^b {
                g_A(x,t;a,\theta) g_A(x,t;\theta,b) 
                - g_A(x,t;a,b) .
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_a^0 {
                g_B(x,t;a,\theta) g_A(x,t;\theta,b)
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_0^b {
                g_A(x,t;a,\theta) g_B(x,t;\theta,b)
            } d\theta,
    \end{aligned}
\end{align}

@@

Each deme has exactly $N_d$ diploid individuals at the start of each generation.
At the start of each generation, every individual migrates to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
Migrants past either end of the range remained at the terminal demes.
Then, 
fitnesses were computed (heterozygotes have fitness $1-s$; homozygotes have fitness 1),
and in each deme $N_d$ pairs of parents are chosen, with replacement,
with probability proportional to their fitness and selfing allowed.
Since migration is not conservative, demes may have no available parents;
in this case, parents are chosen from other demes with probability proportional to fitness
multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme.
The next generation is formed by carrying out meiosis in each parent
and combining the resulting gametes such that each pair of parents leaves one descendant.
Meiosis results in alternating blocks of the gamete's chromsome
being inherited from the two parental chromosomes,
with the blocks separated by a Poisson(1) number of uniformly chosen recombination points along the chromosome,
and the order of the parental chromosomes chosen randomly.
The simulation software works by recording for each chromosome
a list of ancestry breakpoints and the index of the ancestor in the first generation
that the chromosome inherited that region of genome from;
genotyping at individual loci was performed by looking up the ancestor's genotype.

Hybrid zones of ages 10, 100 and 1000 generations \alisa{these numbers may need to be revised?} were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$. From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$). These were compared to the output of neutral simulations with $s=0$

@@

\subsection*{Statistics}
With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci. In reality, however, such loci are not known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection. The most straightforward measure is $l_I(m)$ the length of haplotype chunk that a sampled chunk containing genomic position $b$, conditioned on ancestry $I$ at the site. Based on our expectation of the distribution of block lengths, we compute $l_B(m)$ for $m$ across a chromosome.

@@

Since heterozygotes for the selected allele are only found at high frequency in the cline,
and lineages are generally pushed away from the cline but have no bias far away,
the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
and so the width of the genomic region showing clines about the selected locus could be 
substantially larger than $1/\tau$. However, this distinction appears hard to observe for realistic parameter values.

@@

\subsection*{The distribution of ancestry tract lengths}
The distribution of contiguous ancestry block lengths necessarily contains more information than allele frequency alone. We are interested specifically in the how tracts of ancestry surrounding the selected locus compares to the rest of the genome. As expected, regions surrounding a locus under selection are more resistant to introrgression (Fig~\ref{Fig:resistanceToIntrogression1000g},~\ref{Fig:resistanceToIntrogression100g}).

When present, however, we expect an unfit migrant haplotype to be longer than the genome-wide average, and indeed we observe a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Fig.~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}). This pattern provides a potential avenue for identifying targets of selection by looking for a signal of longer-than-average tracts surrounding a position, compared to a genome-wide background. (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone. A signal is also present for, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).

In the context of our model, we expect that\alisa{, in the absence of selection,} stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
because they have had more opportunities to recombine with $B$ haplotypes. 
However, stretches of $A$ ancestry that contain a selected site are expected to be longer
than those that do not at the same spatial location,
because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
concretely, they move at speed roughly $\sigma/\sqrt{s}$,
so we expect (selected) $A$ alleles at distance $x$ from the cline center
to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago \alisa{(compared to $\sigma$?? for a neutral allele)},
and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
\plr{(check predicted decay of mean length $1/x$ in simulations).}
For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

@@

\subsection*{Lengths of adjacent blocks}
We find that the length of blocks adjacent to blocks of `wrong' ancestry surrounding a selected site are slightly shorter than average (Figure~\ref{Supp:adjacentBlocks}). This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by taking the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).

@@

\begin{figure}
\includegraphics{figs/blocksAlongChromAncBConditioning}
\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$. Here each line represents  $l_B(\mathbf{m})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m_i)$ along the chromosome in the deme.}\label{Fig:blockLengths}
\end{figure}

\subsection*{The number of `unfit' ancestors}
Although the theoretical model presented here does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations. In particular, we wish to know the relative number of ancestors from $\tau$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry. This is a reflection of the average size of a haplotype family. We find that, as distance from the zone center increases, the average family size of an ancestor from time $\tau$ decreases (Figure~\ref{Fig:family_size}). One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection.

@@

Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.

>>>>

\documentclass[11pt,letterpaper]{article}
\usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{color}
\usepackage{amssymb}
\usepackage{setspace}

@@

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.  Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.

Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}. 
%The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

@@

Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. Our goal is to understand how selection against hybrids interacts with dispersal and recombination to produce observed patterns ancestry around selected sites in recently established hybrid zones. In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness. In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and  patterns of introgression can provide information about selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}. 
\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}

@@

We model selection as a simple case of under dominance at a single locus.  Specifically, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes. \yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those generated under complex epistatic systems}.

Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}). 
The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus). 
For every other locus, we keep track of $z$, the identity at $L$, across time and space. If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise. Recombination between two loci is modeled as a poisson process with rate $r$. The identity $z$ for locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.

\begin{figure}
\includegraphics{figs/BM_schematic}
\caption{A cartoon representation of a haplotype under the model. The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day ($\tau=0$). The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A segregates at lower frequency for $x>0$. Here, because of negative selection, $L_B$ migrates rapidly back to $x>0$ from a position $x<0$. Red dots represent recombination events along the entire haplotype. When a recombination separates two chunks of the sampled haplotype takes place, a branching event occurs and these splitting lineages are represented by colored lines (here the initial haplotype is broken up into 3 chunks over 100 generations). Subsequent recombination events can switch the selected background to which a chunk is linked.  Here, yellow lineages are linked to $L_A$ , and blue lineages are linked to $L_B$. Based on position at time $\tau=100$, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
\end{figure}
%As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone. This means that recombination events that occur away from the center of the zone do not result in a change in $a$.

@@

Consider the collection of $A$ alleles found at geographic location $x$.
The expected number, among these alleles, that had a parent at location $y$ in the previous generation is proportional to the number of  $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$. In other words, lineages move as a random walk determined by the dispersal kernel,
but biased towards locations where the selected allele they carry is at higher frequency.

\alisa{Maybe describe, in a few words, how you're constructing the model that is in the Appendix.}
Making the same assumptions as for  equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill \alisa{better word?} on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
Therefore, roughly, lineages on 
then roughly, lineages on ``their own'' side wander randomly,
while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
back towards the side where they are more common. \alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
Lineages of $B$ alleles do the same thing, of course, but in the opposite direction. \alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}

\paragraph{Lineages at linked loci}
The behavior of a lineage at a linked locus is similar to a selected locus,
except that when in heterozygotes the lineage may move between backgrounds:
if we follow back through time the lineage of a locus we find linked to an $A$ allele, 
it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
but if it spends time in the clinal region where recombination may happen,
it may be found to have been inherited from a $B$-carrying individual,
whose ancestors will tend to be more from the right.

@@

\paragraph{PDE for linked clines}
Clines at linked sites are given by
the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
We denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation. \alisa{Is the above equation true? Shouldn't it depend on $r$?}
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            + r (1-p(x,t))(q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} \\
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$. \alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
This can be expressed as differential operators, and written as:

We will have more use for the differential operators on the right-hand sides of these equations,
so write these also as 
$$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$$
and 
$$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$$.
\alisa{A few words about what $G$ is?}

@@

\paragraph{Haplotype lengths}
We can also find the probability of seeing an entire block of genome (haplotype)
of a single ancestry at a given location and time.
To quantify haplotype abundances, we start by considering a segment of the genome between positions $a$ and $b$, relative to the selected site. We are interested in the probability  $g_A(x,t;a,b)$ that the entire segment $(a,b)$ in an individual sampled at spatial location $x$ at time $t$ after the cline has formed is of ancestry $A$.

\paragraph{General haplotype probabilities}
As in \citet{Sedghifar2015}, we need only keep track of both lineages when a recombination occurs within a block (Figure~\ref{Fig:schematic}),
so that lineages will act like a particular class of labeled, branching diffusions,
where the total branching rate is conserved. We will not consider subsequent coalescence.
The general description of the process is as follows:
The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
with recombination distance $r$ equal to the rate at which the segment recombines away from the selected site.
(If the selected site lies inside the segment, $r=0$.)
Additionally, at rate equal to the genetic length of the segment,
a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
each of which proceeds as before, independently. $g_A(x,t;a,b)$ is then the  probability that all branches are found on the same side when the cline formed.

This is similar to equation \eqref{eqn:q_pde} with a term added for branching, so that 
\plr{write this equation more compactly?}
\begin{align}
    \begin{aligned} \label{eqn:g_pde}
        \dot g_A(x,t;a,b) 
            &= G_{r(a,b)}^A g_A(x,t;a,b) 
            - (b-a) g_A(x,t;a,b) 
            \\ {} & \qquad 
            + p(x,t) \int_a^b {
                g_A(x,t;a,\theta) g_A(x,t;\theta,b) 
                - g_A(x,t;a,b) .
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_a^0 {
                g_B(x,t;a,\theta) g_A(x,t;\theta,b)
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_0^b {
                g_A(x,t;a,\theta) g_B(x,t;\theta,b)
            } d\theta,
    \end{aligned}
\end{align}

@@

Each deme has exactly $N_d$ diploid individuals at the start of each generation.
At the start of each generation, every individual migrates to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
Migrants past either end of the range remained at the terminal demes.
Then, 
fitnesses were computed (heterozygotes at the selected locus have fitness $1-s$; homozygotes have fitness 1),
and in each deme $N_d$ pairs of parents are chosen, with replacement,
with probability proportional to their fitness and selfing allowed.
Since migration is not conservative, demes may have no available parents;
in this case, parents are chosen from other demes with probability proportional to fitness
multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme.
The next generation is formed by carrying out meiosis in each parent
and combining the resulting gametes such that each pair of parents leaves one descendant.
Meiosis results in alternating blocks of the gamete's chromsome
being inherited from the two parental chromosomes,
with the blocks separated by a Poisson(1) number of uniformly chosen recombination points along the chromosome,
and the order of the parental chromosomes chosen randomly.
The simulation software works by recording for each chromosome
a list of ancestry breakpoints and the index of the ancestor in the first generation
that the chromosome inherited that region of genome from;
genotyping at individual loci was performed by looking up the ancestor's genotype.

Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$. From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$). These were compared to the output of neutral simulations with $s=0$

@@

\subsection*{Measures of introgression}
With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci. In reality, however, such loci are not known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection. The most straightforward measure is $l_I(m)$ the length of haplotype chunk that a sampled chunk containing genomic position $b$, conditioned on ancestry $I$ at the site. Based on our expectation of the distribution of block lengths, we compute $l_B(m)$ for $m$ across a chromosome.

@@

Since heterozygotes for the selected allele are only found at high frequency in the cline,
and lineages \alisa{on the ``correct" side?} are generally pushed away from the cline but have no bias far away,
the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
and so the width of the genomic region showing clines about the selected locus could be 
substantially larger than $1/\tau$. However, this distinction appears hard to observe for realistic parameter values.

@@

\subsection*{The distribution of ancestry tract lengths}
The distribution of contiguous ancestry block lengths necessarily contains more information than allele frequency alone. We are interested specifically in the how tracts of ancestry surrounding the selected locus compares to the rest of the genome.

As expected, regions surrounding a locus under selection are more resistant to introrgression (Fig~\ref{Fig:resistanceToIntrogression1000g},~\ref{Fig:resistanceToIntrogression100g}).  When present, however, we expect an unfit migrant haplotype to be longer than the genome-wide average, and indeed we observe a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Fig.~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}). This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}). A signal is also present for, but less well resolved when considering $l_\cdot$ (i.e., when ancestry is not taken into account, Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).

In the context of our model, we expect that\alisa{, in the absence of selection,} $l_A$ across the genome will tend to be shorter the further one into the $B$ side of the cline,
because they have had more opportunities to recombine with $B$ haplotypes. 
However, stretches of $A$ ancestry that contain a selected site are expected to be longer
than those that do not at the same spatial location,
because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
concretely, they move at speed roughly $\sigma/\sqrt{s}$,
so we expect (selected) $A$ alleles at distance $x$ from the cline center
to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago \alisa{(compared to $\sigma$?? for a neutral allele)},
and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
\plr{(check predicted decay of mean length $1/x$ in simulations).}
For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

@@

\subsection*{Lengths of adjacent blocks}
We find that $l_A(k\pm)$, the length of blocks adjacent to blocks containing $k$, are shorter than average when $k$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side" of the side of the zone.  (Figure~\ref{Supp:adjacentBlocks}). This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by taking the ratio $\frac{2\sum{l_B(k_i)}}{\sum{l_A(k_i+)+l_A(k_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).

@@

\begin{figure}
\includegraphics{figs/blocksAlongChromAncBConditioning}
\caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$. Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.}\label{Fig:blockLengths}
\end{figure}

\subsection*{The number of `unfit' ancestors}
Although the theoretical model presented here does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations. In particular, we wish to know the relative number of ancestors from $\tau$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry. This is a reflection of the average size of a haplotype family. We find that, as distance from the zone center increases, the average family size of an ancestor from time $\tau$ decreases (Figure~\ref{Fig:family_size}). One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection. This can potentially provide additional signal for characterizing lineages strongly influenced by selection.

@@

%Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.

%\alisa{epistasis}
%\alisa{selection here is STRONG, not sure how well this will work on systems in which selection on any given locus is small (most systems)}
